State of the Practice for Lattice Boltzmann Method Software

### 2022-04-23 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Addressing more passive voice problems

Filtering by scope, usage, and age decreased the size of the list to 23
packages. We removed 22 packages for the following reasons: we cannot install
them, they are incomplete, source code is not publicly available, a license is
needed, or the project is out of scope. \citet{Michalski2021} lists the
eliminated software packages. Of the remaining 23 packages, we kept some despite
their `dead` status because of their presence on authoritative LBM software
lists and because of their surface excellence. After the initial measurement
exercise, which took place in mid-2020, we added one more package in January
2022: Musubi. We found Musubi after our initial search, but we felt obliged to
include it since it clearly fits our criteria. Time constraints meant we didn't
update the previous manual data collection, but we did redo the automated data
collection.

@@

We designed the questions to be unambiguous, quantifiable, and measurable with
limited time and domain knowledge. We categorize them by quality, with an extra
category for summary information (Figure~\ref{measurement_template_image}). The
summary section provides general data, such as the software name, number of
developers, etc. \citet{gewaltig2012quality} provide us with our software
categories. Public means software available to all, while private means software
available only to a specific group.  The concept category defines software
written to demonstrate algorithms or concepts. The three categories of
development models are open source, where source code is freely available under
an open-source license; free-ware, where a binary or executable is provided for
free; and commercial, where the user must pay.

Several of the qualities use the word ``surface'' to highlight that, for these
qualities, the best that we can do is a shallow measure. For instance, we do not
conduct experiments to measure usability directly. Instead, we look for an
indication that usability is considered by searching for cues in the
documentation, such as the presence of getting started instructions, a user
manual or documentation of expected user characteristics.

@@

Since the LBM domain has a reasonably small scope, we can view the software as
constituting a program family.  \citet{parnas1976design} defines a program
family as ``a set of programs whose common properties are so extensive that it
is advantageous to study the common properties of the programs before analyzing
individual members.'' Studying the common properties within a family of related
programs is termed a domain analysis.

The domain analysis (also called a commonality analysis) consists of
understanding the commonalities, variabilities, and common terminology for a
program family \citep{weiss1997defining}. Commonalities are goals, theories,
models, definitions, and assumptions that are common between family members.
Variabilities are goals, theories, models, definitions, and assumptions that
differ between family members. Associated with each variability are parameters
of variation, which summarize the possible values for that variability, along
with their potential binding time. The binding time is when the value of a
variability is set. Variabilities are set at specification time, build time
(when the program compiles) or run time (when the code executes). For the
current assessment, the packages all have the commonality of using LBM
techniques to simulate the motion of fluids. Variabilities for LBM packages
include the following: the dimension of problems they simulate, whether 
(and how) they support parallel computation, whether compressibility can
be considered, whether reflexive boundaries can be employed, whether the
simulation handles multi-fluids, whether the simulation handles turbulence, and
whether complex geometries can be simulated.

To keep the current presentation focused, we only present the results of our
shallow domain analysis. For the LBM domain, a table was constructed that
distinguishes the programs by their variabilities. In research software, the
variabilities are often assumptions. Table~\ref{tbl_features} lists the
variabilities as columns, with the software alphabetically ordered.
\citet{Michalski2021} provides a detailed analysis of the commonalities and
variabilities between LBM packages.

@@

All 24 packages are installable on a Unix-like system. Eight packages are also
installable on Windows, and five on macOS. Twenty packages list their operating
system requirements, but only four (ESPResSo, LUMA, Mechsys, Sailfish) list
compatible operating system versions. Ubuntu was used to test all but one of the
packages. TCLB was tested on CentOS, as suggested by its installation
instructions.

@@

Errors encountered during the installation process were often quickly fixed,
thanks to descriptive error messages. Systems that provided vague error
messages, such as messages that did not specify which action or file was at
fault, were harder to troubleshoot. Only three software packages (HemeLB, LB3D,
lbmpy) that displayed a descriptive error message were not recoverable, and most
of these instances were due to hardware and operating system incompatibility,
such as the requirement of CUDA. Fourteen software packages broke during
installation. Some packages, such as LB2D-Prime and LB3D-Prime did not provide a
definitive message of the success or failure of the installation. In these
instances, validating the installation required performing a tutorial or running
a script, as described below, if these were available.

@@

Seventeen software packages require installation of less than ten dependencies.
All but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Twenty of the software packages do not explicitly indicate the versions of
software dependencies. Some software package installation issues, specifically
those occurring when the manual installation of dependencies is required, might
be avoided if versions of dependencies are specified. Sixteen software packages
do not have detailed instructions for installing dependencies. Sixteen software
packages have less than ten manual installation steps. If one command installs
dependencies, then none of the software packages take more than 20 steps to
install. The average number of steps is eight, and the fewest is two
(LB3D-Prime).

@@

Seventeen software packages explicitly reference domain theory, but a complete
and rigorous requirements specification, in the software engineering sense
\citep{IEEE1998, RobertsonAndRobertson1999Vol, ESA1991},  is absent from all
projects. Software packages that include a subset of a requirements
specification, such as DL\_MESO (LBE), keep the information brief and include it
within other documents. When present, theory related information is generally
found within a user manual, on a webpage, or in a cited publication. In the
latter case tracking down the knowledge may take a significant time.

Document generation tools are explicitly used by 13 software packages. Sphinx is
used by eight of them, and Doxygen is used by eight. Several of the packages use
both.

Tutorials are available for 19 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.  A
particularly  detailed tutorial that includes expected output is provided by
\href{https://www.walberla.net/doxygen/index.html} {waLBerla}.

@@

The use of Continuous Integration (CI) tools and techniques alludes to a more
refined development process where faults are isolated and better recognized.
Only three of the packages (ESPResSo, Ludwig, Musubi) mentioned applying the
practice of CI in their development process.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the ranking for surface
correctness and verifiability. Software packages with a higher score tend to
have theory documentation. They also explicitly use at least one document
generation tool that helps with verification, thus building confidence in
correctness. The top ranked software packages all include getting started
tutorials, and most of these include expected output. The two top ranked
packages, Ludwig and Musubi, indicate the use of unit testing in their
documentation. They also incorporated CI in the development process.
Furthermore, eight of the top 10 ranked packages are noted as being alive.

@@

The analysis of surface reliability focused on package installation and
tutorials. Errors occurred when installing 16 of the software packages. Every
instance prompted an error message. These messages indicated unrecognized
commands (even when following the installation guide), missing links, missing
dependencies, and syntax errors in code files. In some instances the error
messages were vague. Several automatic installation processes could not find or
load dependencies. In these instances the installation tried to access outdated
external repositories. Seven of the installations were recovered and verified,
and one of the installations (LB3D-Prime) was assumed to be recovered due to the
absence of any way to verify otherwise. The installation of eight of the software
packages could not be recovered. Most of these broken installations could not
find external dependencies, encountered system incompatibilities, or displayed
vague error messages.

Of the 14 software packages that installed correctly and also have tutorials,
four (pyLBM, \esp, LIMBES, Ludwig) broke during tutorial testing. All of
these instances resulted in an error message being displayed. One error (pyLBM)
was due to a missing tutorial dependency, another (Ludwig) was due to an invalid
command despite following the tutorial, and the final two errors were vague
execution errors. Of the four broken tutorial instances, only the one due to a
missing dependency was recoverable.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. All the top
five ranked packages have tutorials. The package pyLBM broke during tutorial
testing, but a descriptive error message helped in recovery. Nine of the top 10
ranked packages are noted as being alive.

@@

The software packages were tested for handling unexpected input, including
incorrect data types, empty input, and missing files or links. Success is
predicated on a reasonable response from the system, including appropriate error
messages and an absence of unrecoverable system failures.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All the software packages that installed
correctly passed this test. They output descriptive error messages or at least
did not crash. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
runtime robustness. All successfully installed software packages that require a
plain text input file correctly handled an unexpected change to these input
files, including a replacement of new lines with carriage returns. Furthermore,
nine of the top 10 ranked packages are noted as being alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of run-time performance is
not appropriate. Instead, we looked through each software package's artifacts for
evidence that performance was considered. The artifacts of 18 software packages
mentioned parallelization. This included GPU processing and the CUDA parallel
computing platform, which were mentioned in the artifacts of six packages
(ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). When a high degree of
parallelization is possible GPUs provide superior speed compared to CPUs. The
software package TCLB is implemented in a highly efficient multi-GPU code to
achieve performance suitable for model optimization \citep{rutkowski2020open}.
In the Ludwig package, a so-called mixed mode approach is used where
fine-grained parallelism is implemented on the GPU, and MPI is used for even
larger scale parallelism \citep{gray2013ludwig}. While one software package
(Sailfish) required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce,
pyLBM, TCLB) have the option of using either the GPU or the CPU. In general, the
packages that require GPU and CUDA have better performance at the expense of
installability and surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 19
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output. Most
user manuals are in the form of a file that can be downloaded, while some are
rendered on a webpage. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their webpages. Expected
user characteristics are documented in five software packages (laboetie, LIMBES,
Ludwig, Musubi, Palabos). Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All but one
of the packages (LIMBES) have a user support model, and many of them have
multiple avenues of user support. The most popular avenue of support is Git,
followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Figure~\ref{Fig_Usability} shows the surface usability ranking of the software
packages. Software packages with a high score have a tutorial and user manual,
sometimes have documented user characteristics, and have at least one user
support model. Many packages have several user support models. Furthermore, four
of the top five packages are noted as being alive.

@@

Software packages were reviewed for the presence of artifacts. Every type of
artifact or file that is not a code file was recorded. (The list of artifacts
will be discussed further in Section~\ref{Sec_CompareArtifacts}.)  The software
packages were also reviewed for software release and documentation version
numbers. This information can be used to troubleshoot issues and organize
documentation. All but three software packages (LatBo.jl, LB3D-Prime, MechSys)
have source code release and documentation version numbers.

When present, information on how code is reviewed, or how to contribute to the
project was also noted. In total, 11 software packages have this information,
which was found in various artifacts, including in developer guides, contributor
guides, user guides, developer webpages, and README files.

Issue tracking is used in 23 software packages, 15 of which use GitHub or GitLab
to host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50 percent of their issues closed. All the top five overall ranked
packages have most of their issues closed. Alive packages (11 use issue
tracking) have 64\% of their issues closed, while dead packages (3 use issue
tracking) have 71\% of their issues closed. This information is presented in
Table~\ref{gitrepodata}. With respect to the host for the git repository, 13
packages use GitHub and two (Palabos, waLBerla) use GitLab. Of the other
packages, one package (SunlightLB) uses CVS for issue tracking and version
control, and seven packages do not appear to use any issue tracking system.

@@

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}, sorted
by the percentage. Packages with a higher percentage of comments were assumed to
be more maintainable. Comments represent more than 10 percent of code files in
16 packages, and the average percentage of code comments is about 14 percent.
All the top five overall ranked packages have more than the average. LUMA has
only 0.2 percent comments, the fewest of any package. This package has the most
lines of source code, with over four million. The next largest package is
\esp{} with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to manage their issue tracking, having most of their issues
closed. Their code files are well commented with more than 10 percent of the
code being comments. Four of the top five ranked packages are noted as being alive.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages. Software packages with a high score have thousands of source code
files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modelling. For this reason, a comparison with other software packages is
not on a level field. However, these packages do have an abundance of reusable
components. Nine of the top 10 ranked packages are noted as being alive.

@@

Figure~\ref{Fig_Understandability} shows the surface understandability ranking.
Software packages with a high score have a consistent indentation and formatting
style, and consistent, distinctive, and meaningful code identifiers. They also
have symbolic constants, and explicitly identify mathematical and LBM
algorithms. Their comments are clear and indicate what is being done in the
source code. The source code is well modularized and structured. All the top
five ranked packages are noted as being alive.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also have
detailed and easy to access notes accompanying software releases. Four of the
top five ranked packages are noted as being alive.

>>>>

Filtering by scope, usage, and age decreased the size of the list to 23
packages. We removed 22 packages for the following reasons: we cannot install
them, they are incomplete, source code is not publicly available, we are not
licensed to use them, or the project is out of scope. \citet{Michalski2021}
lists the eliminated software packages. Of the remaining 23 packages, we kept
some despite their `dead` status because of their presence on authoritative LBM
software lists and because of their surface excellence. After the initial
measurement exercise, which took place in mid-2020, we added one more package in
January 2022: Musubi. We found Musubi after our initial search, but we felt
obliged to include it since it clearly fits our criteria. Time constraints meant
we didn't update the previous manual data collection, but we did redo the
automated data collection.

@@

We designed the questions to be unambiguous, quantifiable, and measurable with
limited time and domain knowledge. We categorize them by quality, with an extra
category for summary information (Figure~\ref{measurement_template_image}). The
summary section provides general data, such as the software name, number of
developers, etc. \citet{gewaltig2012quality} provide us with our software
categories. Public means software available to all, while private means software
available only to a specific group.  The concept category defines software
written to demonstrate algorithms or concepts. The three categories of
development models are open source, where source code is freely available under
an open-source license; free-ware, which freely provides a binary or executable,
but not source code; and commercial, where the user must pay for the software.

Several of the qualities use the word ``surface'' to highlight that, for these
qualities, the best that we can do is a shallow measure. For instance, we do not
conduct experiments to measure usability directly. Instead, we look for
usability considerations by searching for cues in the documentation, such as the
presence of getting started instructions, a user manual or documentation of
expected user characteristics.

@@

Since the LBM domain has a reasonably small scope, we can view the software as
constituting a program family.  \citet{parnas1976design} defines a program
family as ``a set of programs whose common properties are so extensive that it
is advantageous to study the common properties of the programs before analyzing
individual members.'' Domain analysis is the term for studying the common
properties within a family of related programs.

The domain analysis (also called a commonality analysis) consists of
understanding the commonalities, variabilities, and common terminology for a
program family \citep{weiss1997defining}. Commonalities are goals, theories,
models, definitions, and assumptions that are common between family members.
Variabilities are goals, theories, models, definitions, and assumptions that
differ between family members. Associated with each variability are parameters
of variation, which summarize the possible values for that variability, along
with their potential binding time. The binding time is when we specify a value
for a variability. Binding time options include specification time, build time
(when the program compiles) and run time (when the code executes). For the
current assessment, the packages all have the commonality of using LBM
techniques to simulate the motion of fluids. Variabilities for LBM packages
include the following: the dimension of problems they simulate, whether (and
how) they support parallel computation, whether compressibility can be
considered, whether reflexive boundaries can be employed, whether the simulation
handles multi-fluids, whether the simulation handles turbulence, and whether
complex geometries can be simulated.

To keep the current presentation focused, we only present the results of our
shallow domain analysis. We distinguish programs by their variabilities, which
for research software are often assumptions. Table~\ref{tbl_features} lists the
variabilities as columns, with the software alphabetically ordered.
\citet{Michalski2021} provides a detailed analysis of the commonalities and
variabilities between LBM packages.

@@

All 24 packages are installable on a Unix-like system. Eight packages are also
installable on Windows, and five on macOS. Twenty packages list their operating
system requirements, but only four (ESPResSo, LUMA, Mechsys, Sailfish) list
compatible operating system versions. Ubuntu provided the testing platform for
all but one of the packages. TCLB was tested on CentOS, as suggested by its
installation instructions.

@@

We quickly fixed errors encountered during the installation process, thanks to
descriptive error messages. Systems that provided vague error messages, such as
messages that did not specify which action or file was at fault, were harder to
troubleshoot. Only three software packages (HemeLB, LB3D, lbmpy) that displayed
a descriptive error message were not recoverable, and most of these instances
were due to hardware and operating system incompatibility, such as the
requirement of CUDA. Fourteen software packages broke during installation. Some
packages, such as LB2D-Prime and LB3D-Prime did not provide a definitive message
of the success or failure of the installation. In these instances, validating
the installation required performing a tutorial or running a script, as
described below, if these were available.

@@

Seventeen software packages require installation of less than ten dependencies.
All but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Twenty of the software packages do not explicitly indicate the versions of
software dependencies. Some software package installation issues could have been
avoided by providing a list of dependencies. Sixteen software packages do not
have detailed instructions for installing dependencies. Sixteen software
packages have less than ten manual installation steps. If one command installs
dependencies, then none of the software packages take more than 20 steps to
install. The average number of steps is eight, and the fewest is two
(LB3D-Prime).

@@

Seventeen software packages explicitly reference domain theory, but a complete
and rigorous requirements specification, in the software engineering sense
\citep{IEEE1998, RobertsonAndRobertson1999Vol, ESA1991},  is absent from all
projects. Software packages that include a subset of a requirements
specification, such as DL\_MESO (LBE), keep the information brief and include it
within other documents, such as a user manual, webpage, or a publication. In the
latter case tracking down the knowledge may take significant time.

Thirteen software packages explicitly use document generation tools. Eight of
them use Sphinx and the same number use Doxygen. Several packages use both.

Tutorials are available for 19 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.
\href{https://www.walberla.net/doxygen/index.html} {waLBerla} provides a
particularly detailed tutorial, including expected output.

@@

The use of Continuous Integration (CI) tools and techniques alludes to a refined
development process that quickly isolates and fixes faults. Only three of the
packages (ESPResSo, Ludwig, Musubi) mentioned applying the practice of CI in
their development process.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the ranking for surface
correctness and verifiability. Software packages with a higher score tend to
have theory documentation. They also explicitly use at least one document
generation tool that helps with verification, thus building confidence in
correctness. The top ranked software packages all include getting started
tutorials, and most of these include expected output. The two top ranked
packages, Ludwig and Musubi, indicate the use of unit testing in their
documentation. They also incorporated CI in the development process.
Furthermore, eight of the top 10 ranked packages are alive.

@@

The analysis of surface reliability focused on package installation and
tutorials. Errors occurred when installing 16 of the software packages. Every
instance prompted an error message. These messages indicated unrecognized
commands (even when following the installation guide), missing links, missing
dependencies, and syntax errors in code files. In some instances the error
messages were vague. Several automatic installation processes could not find or
load dependencies. In these instances the installation tried to access outdated
external repositories. We recovered and verified seven of the installations,
and we assumed that one of the installations (LB3D-Prime) was recovered due to the
absence of any way to verify otherwise. The installation of eight of the software
packages could not be recovered. Most of these broken installations could not
find external dependencies, encountered system incompatibilities, or displayed
vague error messages.

Of the 14 software packages that installed correctly and also have tutorials,
four (pyLBM, \esp, LIMBES, Ludwig) broke during tutorial testing. All of these
instances resulted in the display of an error message. One error (pyLBM) was due
to a missing tutorial dependency, another (Ludwig) was due to an invalid command
despite following the tutorial, and the final two errors were vague execution
errors. Of the four broken tutorial instances, only the one due to a missing
dependency was recoverable.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. All the top
five ranked packages have tutorials. The package pyLBM broke during tutorial
testing, but a descriptive error message helped in recovery. Nine of the top 10
ranked packages are alive.

@@

Testing includes handling unexpected input, such as incorrect data types, empty
input, and missing files or links. A reasonable response, including appropriate
error messages and an absence of unrecoverable system failures, means success.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All the software packages that installed
correctly passed this test. They output descriptive error messages or at least
did not crash. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
runtime robustness. All successfully installed software packages that require a
plain text input file correctly handled an unexpected change to these input
files, including a replacement of new lines with carriage returns. Furthermore,
nine of the top 10 ranked packages are alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of run-time performance is
not appropriate. Instead, we looked through each software package's artifacts
for evidence of considering performance. The artifacts of 18 software packages
mentioned parallelization. This included GPU processing and the CUDA parallel
computing platform, which were mentioned in the artifacts of six packages
(ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). When a high degree of
parallelization is possible, GPUs provide superior speed compared to CPUs. The
software package TCLB is implemented in a highly efficient multi-GPU code to
achieve performance suitable for model optimization \citep{rutkowski2020open}.
In the Ludwig package uses a so-called mixed mode approach implementing
fine-grained parallelism on the GPU.  MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. In general, the packages
that require GPU and CUDA have better performance at the expense of
installability and surface reliability.

@@

We reviewed software package artifacts for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 19
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output. Most
user manuals are in the form of a file that can be downloaded, while some are
rendered on a webpage. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their webpages. Five
software packages (laboetie, LIMBES, Ludwig, Musubi, Palabos) document expected
user characteristics. Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All but one
of the packages (LIMBES) have a user support model, and many of them have
multiple avenues of user support. The most popular avenue of support is Git,
followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Figure~\ref{Fig_Usability} shows the surface usability ranking of the software
packages. Software packages with a high score have a tutorial and user manual,
sometimes have documented user characteristics, and have at least one user
support model. Many packages have several user support models. Furthermore, four
of the top five packages are alive.

@@

We reviewed software packages for the presence of artifacts and recorded every
type of artifact that is not a code file. (The list of artifacts will be
discussed further in Section~\ref{Sec_CompareArtifacts}.)  We also reviewed the
software packages for software release and documentation version numbers. This
information can be used to troubleshoot issues and organize documentation. All
but three software packages (LatBo.jl, LB3D-Prime, MechSys) have source code
release and documentation version numbers.

When present, we noted how the team reviews code, and how users can contribute
to the project. In total, 11 software packages have this information, which was
found in various artifacts, including in developer guides, contributor guides,
user guides, developer webpages, and README files.

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50 percent of their issues closed. All the top five overall ranked packages
have most of their issues closed. Alive packages (11 use issue tracking) have
64\% of their issues closed, while dead packages (3 use issue tracking) have
71\% of their issues closed. Table~\ref{gitrepodata} presents this information.
With respect to the host for the git repository, 13 packages use GitHub and two
(Palabos, waLBerla) use GitLab. Of the other packages, one package (SunlightLB)
uses CVS for issue tracking and version control, and seven packages do not
appear to use any issue tracking system.

@@

Table~\ref{gitrepodata} shows the percentage of code that is comments, in
decreasing order. Maintainability improves for packages with a
higher percentage of comments. Comments represent more than 10 percent of code
files in 16 packages, and the average percentage of code comments is about 14
percent. All the top five overall ranked packages have more than the average.
LUMA has only 0.2 percent comments, the fewest of any package. This package has
the most lines of source code, with over four million. The next largest package
is \esp{} with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to manage their issue tracking, having most of their issues
closed. Maintainable packages by our ranking have more than 10 percent of the
code as comments. Four of the top five ranked packages are alive.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages. Software packages with a high score have thousands of source code
files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modelling. For this reason, a comparison with other software packages is
not on a level field. However, these packages do have an abundance of reusable
components. Nine of the top 10 ranked packages are alive.

@@

Figure~\ref{Fig_Understandability} shows the surface understandability ranking.
Software packages with a high score have a consistent indentation and formatting
style, and consistent, distinctive, and meaningful code identifiers. They also
have symbolic constants, and explicitly identify mathematical and LBM
algorithms. Their comments are clear and indicate what is being done in the
source code. The source code is well modularized and structured. All the top
five ranked packages are alive.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also have
detailed and easy to access notes accompanying software releases. Four of the
top five ranked packages are alive.
